---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Alberte Seeberg"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia, focusing on pitch.
In the course of this assignment we will use them to try to automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.

### Question 1
Build a logistic regression to see whether you can diagnose schizophrenia from pitch range only.

Calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve) on a logistic regression using the full dataset. Don't forget the random effects!

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures.

N.B. the predict() function generates probabilities (the full scale between 0 and 1). A probability > .5 indicates a choice of 1, below a choice of 0.
N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?

```{r}
#Setting working directory
setwd("~/Desktop/3. Semester/Experimental methods 3/Portfolio 3/Portfolio 3")

#Reading libraries
library(lmerTest)
library(caret)
library(pROC)

#Reading the data
AllData=read.csv("AllData.csv")

AllData=AllData[-1]

AllData$Subject=as.factor(AllData$Subject)

#Making a generalized linear model
model1=glmer(Diagnosis~range + (1+Trial|Subject), AllData, family="binomial")
summary(model1)

#Creating confusion matrix
AllData$PredictionsPerc=predict(model1)
AllData$Predictions[AllData$PredictionsPerc>0]="Schizophrenia"
AllData$Predictions[AllData$PredictionsPerc<=0]="Control"
c=confusionMatrix(data = AllData$Predictions, reference = AllData$Diagnosis, positive = "Schizophrenia") 


#Making a reciever operator curve
rocCurve=roc(response=AllData$Diagnosis, predictor=AllData$PredictionsPerc)
auc(rocCurve)
ci(rocCurve)
plot(rocCurve, legacy.axes=TRUE)

#Cross-validating model

# Adding fold_value
fold_function = function(data, col) {

  data$folds_ID = as.numeric(interaction(data[,col]))
  
  return(data[order(data$folds_ID),])
}

Data_Schizo = subset(AllData, AllData$Diagnosis == "Schizophrenia")
Data_Control = subset(AllData, AllData$Diagnosis == "Control")

Data_Schizo=fold_function(Data_Schizo, "Subject")
Data_Control= fold_function(Data_Control, "Subject")

# Creating folds
folds_S = createFolds(unique(Data_Schizo$folds_ID), 5)
folds_C= createFolds(unique(Data_Control$folds_ID), 5)

# Combining data
fold_list = c()
fold_list$Fold1 = c(folds_S$Fold1, folds_C$Fold1)
fold_list$Fold2 = c(folds_S$Fold2, folds_C$Fold2)
fold_list$Fold3 = c(folds_S$Fold3, folds_C$Fold3)
fold_list$Fold4 = c(folds_S$Fold4, folds_C$Fold4)
fold_list$Fold5 = c(folds_S$Fold5, folds_C$Fold5)

NewDATA = rbind(Data_Schizo, Data_Control)

#Creating a loop

Accuracy=NULL
Specificity=NULL
Sensitivity=NULL
PPV=NULL
NPV=NULL
AUC=NULL

n=1

for (fold in fold_list){
  #Creating a second training dataset 
  train=subset(NewDATA,! (folds_ID %in% fold))
  
  #Creaing a second test dataset
  test=subset(NewDATA, (folds_ID %in% fold))
  
  #Creating a model - train
  TrainModel=glmer(Diagnosis~range + (1+Trial|Subject), train, family="binomial")
  
  test$PredictionsPerc=predict(TrainModel, test, allow.new.levels=TRUE)
  test$Predictions[test$PredictionsPerc>0]="Schizophrenia"
  test$Predictions[test$PredictionsPerc<=0]="Control"
  
  cm=confusionMatrix(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 
  
  Accuracy[n]=cm$overall["Accuracy"]
  
  test$Predictions=as.factor(test$Predictions)
  
  Specificity[n]= specificity(data = test$Predictions, reference = test$Diagnosis, negative = "Control") 
  Sensitivity[n]= sensitivity(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia")
  PPV[n]= posPredValue(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia")
  NPV[n]= negPredValue(data = test$Predictions, reference = test$Diagnosis, negative = "Control") 
  
  rocCurve=roc(response=test$Diagnosis, predictor=test$PredictionsPerc)
  AUC[n]=auc(rocCurve)
    
  n=n+1
 
}

TestData=data.frame(Accuracy, Specificity, Sensitivity, PPV, NPV, AUC)

```

### Question 2

Which single predictor is the best predictor of diagnosis?


```{r}
# Creating a function to find out which single acoustic predictor is the best predictor
CVfunction = function(data, model) {

# Creating empty variables
Accuracy = NULL
Specificity = NULL
Sensitivity = NULL
PPV = NULL
NPV = NULL
AUC = NULL

n = 1

# Creating a loop
for (fold in fold_list){
  
  # Creating training data set
  train = subset(data,! (folds_ID %in% fold))  
  
  # Creating a test data set
  test = subset(data, (folds_ID %in% fold)) 
  
  # Training model
  m1=glmer(model, train, family = "binomial")
  
  # Evaluating model
  test$PredictionsPerc=predict(m1, test, allow.new.levels = TRUE) 
  test$Predictions[test$PredictionsPerc>0]="Schizophrenia" 
  test$Predictions[test$PredictionsPerc<=0]="Control"
  
  cm=confusionMatrix(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 
  
  Accuracy[n] = cm$overall["Accuracy"]
  
  test$Predictions = as.factor(test$Predictions)

  Sensitivity[n] =  sensitivity(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia" )
  
  Specificity [n] = specificity(data = test$Predictions, reference = test$Diagnosis, negative = "Control" ) 

  PPV[n] = posPredValue(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 

  NPV[n] = negPredValue(data = test$Predictions, reference = test$Diagnosis, negative = "Control")  
  
  rocCurve = roc(response = test$Diagnosis,   predictor = test$PredictionsPerc) 
  AUC[n]=auc(rocCurve) 
    
  n=n+1  
}

results = data.frame(Accuracy, Sensitivity, Specificity, PPV, NPV, AUC)

return(results)

}

MeanResults = CVfunction(NewDATA, Diagnosis ~ mean + (1+Trial|Subject))
SDresults = CVfunction(NewDATA, Diagnosis ~ SD + (1+Trial|Subject))
MedianResults = CVfunction(NewDATA, Diagnosis ~ median + (1+Trial|Subject))
RRresults = CVfunction(NewDATA, Diagnosis ~ RR + (1+Trial|Subject))
ENTRresults = CVfunction(NewDATA, Diagnosis ~ ENTR + (1+Trial|Subject))
rENTRresults = CVfunction(NewDATA, Diagnosis ~ rENTR + (1+Trial|Subject))

# Obtaining the mean area under the curve
mean(SDresults$AUC)
# 0.61
mean(MeanResults$AUC) 
# 0.57
mean(MedianResults$AUC) 
# 0.57
mean(RRresults$AUC)
# 0.51
mean(ENTRresults$AUC) 
# 0.61
mean(rENTRresults$AUC) 
# 0.56

```

### Question 3

Now it's time to go wild! Use all (voice-related) variables and interactions you can think of. Compare models and select the best performing model you can find.

Remember:

Out-of-sample error crucial to build the best model!
After choosing the model, send Celine and Riccardo the code of your model

```{r}

#Trying different models

M1=CVfunction(NewDATA, Diagnosis ~ mean*range + (1+Trial|Subject))
M2=CVfunction(NewDATA, Diagnosis ~ SD+ENTR + (1+Trial|Subject))
M3=CVfunction(NewDATA, Diagnosis ~ SD*ENTR + (1+Trial|Subject))
M4=CVfunction(NewDATA, Diagnosis ~ range*L + (1+Trial|Subject))
M5=CVfunction(NewDATA, Diagnosis ~ mean+range + (1+Trial|Subject))
M6=CVfunction(NewDATA, Diagnosis ~ range+L + (1+Trial|Subject))
M7=CVfunction(NewDATA, Diagnosis ~ MAD*SD + (1+Trial|Subject))
M8=CVfunction(NewDATA, Diagnosis ~ MAD+SD + (1+Trial|Subject))
  
  
mean(M1$AUC)
#0.6098
mean(M2$AUC)
#0.6196
mean(M3$AUC)
#0.6161
mean(M4$AUC)
#0.5815
mean(M5$AUC)
#0.6088
mean(M6$AUC)
#0.6097
mean(M7$AUC) #best model
#0.6387
mean(M8$AUC)
#0.5545
```

### Question 4: Report the results

METHODS SECTION: how did you analyse the data?

RESULTS SECTION: can you diagnose schizophrenia based on voice? which features are used? Comment on the difference between the different performance measures.

### Bonus question 5

You have some additional bonus data involving speech rate, pauses, etc. Include them in your analysis. Do they improve classification?

### Bonus question 6

Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them.
